---
title: "Roberts et al. Replication"
author: "Kaylynn Driggers"
date: "2026-03-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lda) #A function used to run LDA in the original paper simulations. 
library(slam)
#install.packages("remotes")
#install.packages("stm")
#install.packages("~/stmdev/rcode/stm_0.03.07.tar.gz", type="source", repos=NULL) #I attempted to recover the early STM version referenced in the original materials,
#but that package version would not install/load in a usable way on my machine.
library(stm) #STM is the package that estimates structured topic models
library(tidyverse) #used for data wrangling, plotting, workflow management. 
knitr::opts_knit$set(root.dir = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Replication Project/Original Replication Data") #Set the working directory to the original replication files so that all data reads point to the same folder used throughout this document.


#These helper functions were referenced in the original STM workflow, but I could not recover
#them in a usable form, so later figures were recreated with current STM tools or custom functions.
#source("~/stmdev/rcode/stm/R/ploteffect.R")
#source("~/stmdev/rcode/stm/R/findthoughts.R")
#source("~/stmdev/rcode/stm/R/plottopics.R")
#source("~/stmdev/rcode/stm/R/plot.stm.R")
#source("~/stmdev/rcode/stm/R/plotPerspectives.R")

```
#Gadarian Experiment Set


#Cleaning Data for Gadarian Data
```{r}
#Read the original Gadarian experiment document-term matrix, vocabulary, and metadata.
dataldac <- readLdac("gadarian_final_tdm.csv") # Reads the text data in LDAC format, which stores each document as term-frequency pairs.
vocabcoarse <- read.csv("gadarian_vocab.csv")$term #Reads the vocabulary file and keep only the term column so topic outputs can be translated back into words.
meta <- read.csv("gadarian_metadata.csv") # Read the document-level metadata used for prevalence covariates and interpretation.
meta$treatment <- as.numeric(meta$treat=="1. worried") # Recode the treatment indicator into a binary variable used in the STM prevalence formulas.

data <- prepDocuments(dataldac, vocabcoarse, meta=meta) # Clean and align documents, vocabulary, and metadata so the STM package can estimate the model without mismatches.
documents <- data$documents # Extract the processed document list used by stm.
vocab <- data$vocab # Extract the cleaned vocabulary that matches the processed documents.
meta <- data$meta # Replace the original metadata with the cleaned metadata returned by prepDocuments.
```


#Selecting the Gadarian STM used for Replication
```{r, include=FALSE}
#In this chunk I did multiple STM runs to find the model that was as close to the original as possible. I followed the author's procedure and did 50 runs. 
#Search and Select Model
set.seed(9432) #set a random seed, after attempting to use the author's seed, just to see if it would work. 
models <- selectModel(documents, vocab, 3, prevalence=~meta$treatment + meta$pid_rep + meta$treatment*meta$pid_rep, runs=50) # Estimate multiple candidate 3-topic STM models and compare them, using treatment, party ID, and their interaction as prevalence covariates.
plotModels(models)  #plotModels() lets me compare candidate solutions across semantic coherence/exclusivity,
# before choosing the run that is most substantively similar to the published results.


#I selected run 4 because its topic structure and downstream effect plots were 
#the closest match to the published figures in the paper, even though the 
#replication did not recover the exact same solution as the original.

mod <- models$runout[[4]] #Select one candidate model from the search output for downstream analysis and figure replication, 4 was closest yet
labelTopics(mod) # inspect the words associated with each recovered topic. 
```



#Figure 2: Simulation 1, Estimated Average Treatment Effect with 95% CI

##Figure 3: Simulation 2, Comparison with LDA

Figures 2 and 3 were simulations run by the authors. The code for how the authors set up and ran these was not included in the replication package. 

#Figure 4: Vocabulary Associated with Topics
```{r}
# Figure 4 displays the main vocabulary associated with the recovered topics,
# while Figures 5 and 6 use representative quotes to illustrate the substance of those topics.

# Topics to display
topics_show <- c(1, 2)

labs <- labelTopics(mod, topics = topics_show, n = 15) # Pull the top words for the selected topics so the figure can show the topic vocabulary.

thoughts <- findThoughts(  # Retrieve exemplar documents that load highly on each topic so the replication can show representative quotes.
  model = mod,
  texts = meta$open.ended.response,
  topics = topics_show,
  n = 5
)
thoughts$docs[[2]]  
# Pick one exemplar per topic.
# Change [X] if another candidate reads better.
quote_t1 <- thoughts$docs[[1]][3]
quote_t2 <- thoughts$docs[[2]][3]


# Custom function used to recreate the topic-word panel layout from the paper.
# I wrote this manually because the original helper plotting functions were unavailable,
# so this function is for reproducing the published figure style rather than estimating the STM itself.

make_stm_topic_panel <- function(
    model,
    topic1,
    topic2 = NULL,
    n = 7,
    frexweight = 0.5,
    wrap_width = 42,
    rect_xleft = 0.08,
    rect_ybottom = 0.12,
    rect_xright = 0.92,
    rect_ytop = 0.88,
    padding_x = 0.05,
    padding_y = 0.06,
    sep_gap = 0.05,
    cex_text = 1.0,
    lineheight = 1.15,
    family = "sans",
    border_lwd = 2,
    sep_lwd = 1.5,
    sep_lty = 2,
    topic_prefix = "Topic",
    return_recorded = FALSE
) {
  if (length(wrap_width) == 1L) {
    wrap_width <- c(wrap_width, wrap_width)
  }
  if (length(wrap_width) != 2L) {
    stop("wrap_width must have length 1 or 2.")
  }
  
  topics_requested <- c(topic1, if (!is.null(topic2)) topic2)
  
  K <- if (!is.null(model$settings$dim$K)) {
    model$settings$dim$K
  } else if (!is.null(model$theta)) {
    ncol(model$theta)
  } else {
    stop("Could not determine the number of topics in the STM object.")
  }
  
  if (any(!is.finite(topics_requested)) ||
      any(topics_requested < 1) ||
      any(topics_requested > K)) {
    stop(sprintf("Topic numbers must be between 1 and %d.", K))
  }
  
  labs <- stm::labelTopics(
    model,
    topics = unique(topics_requested),
    n = n,
    frexweight = frexweight
  )
  
  make_topic_block <- function(topic_id, width) {
    row_id <- match(topic_id, labs$topicnums)
    if (is.na(row_id)) {
      stop(sprintf("Could not find topic %s in labelTopics() output.", topic_id))
    }
    
    if (all(c("prob", "frex", "lift", "score") %in% names(labs))) {
      lines <- c(
        sprintf("%s %d", topic_prefix, topic_id),
       paste(labs$prob[row_id, ], collapse = ", "),
       paste(labs$frex[row_id, ], collapse = ", "),
       paste(labs$lift[row_id, ], collapse = ", "),
      paste(labs$score[row_id, ], collapse = ", "))
    } else if ("topics" %in% names(labs)) {
      # Fallback for content-covariate/aspect models
      lines <- c(
        sprintf("%s %d", topic_prefix, topic_id),
        paste0("Topic Words: ", paste(labs$topics[row_id, ], collapse = ", "))
      )
    } else {
      stop("Unsupported labelTopics() output structure.")
    }
    
    wrapped <- c(
      lines[1],
      unlist(lapply(lines[-1], function(z) strwrap(z, width = width, exdent = 2)))
    )
    
    paste(wrapped, collapse = "\n")
  }
  
  block1 <- make_topic_block(topic1, wrap_width[1])
  block2 <- if (!is.null(topic2)) make_topic_block(topic2, wrap_width[2]) else NULL
  
  op <- par(
    mar = c(0, 0, 0, 0),
    xpd = NA,
    family = family,
    lheight = lineheight
  )
  on.exit(par(op), add = TRUE)
  
  plot.new()
  plot.window(xlim = c(0, 1), ylim = c(0, 1), xaxs = "i", yaxs = "i")
  
  rect(
    rect_xleft, rect_ybottom, rect_xright, rect_ytop,
    col = "white", border = "black", lwd = border_lwd
  )
  
  panel_w <- rect_xright - rect_xleft
  panel_h <- rect_ytop - rect_ybottom
  
  x_left   <- rect_xleft   + padding_x * panel_w
  x_right  <- rect_xright  - padding_x * panel_w
  y_bottom <- rect_ybottom + padding_y * panel_h
  y_top    <- rect_ytop    - padding_y * panel_h
  x_mid    <- (x_left + x_right) / 2
  
  if (is.null(block2)) {
    text(
      x = x_mid,
      y = (y_bottom + y_top) / 2,
      labels = block1,
      adj = c(0.5, 0.5),
      cex = cex_text,
      family = family,
      col = "black"
    )
    
    if (return_recorded) {
      return(recordPlot())
    } else {
      return(invisible(NULL))
    }
  }
  
  h1 <- strheight(block1, units = "user", cex = cex_text, family = family)
  h2 <- strheight(block2, units = "user", cex = cex_text, family = family)
  
  if (!is.finite(h1) || h1 <= 0) h1 <- 1
  if (!is.finite(h2) || h2 <= 0) h2 <- 1
  
  sep_gap_user <- sep_gap * panel_h
  inner_h <- y_top - y_bottom
  usable_h <- inner_h - sep_gap_user
  
  if (usable_h <= 0) {
    stop("Not enough vertical space after padding and separator gap.")
  }
  
  share1 <- h1 / (h1 + h2)
  share2 <- h2 / (h1 + h2)
  
  top_h <- usable_h * share1
  bottom_h <- usable_h * share2
  y_sep <- y_bottom + bottom_h + sep_gap_user / 2
  
  text(
    x = x_mid,
    y = y_top - top_h / 2,
    labels = block1,
    adj = c(0.5, 0.5),
    cex = cex_text,
    family = family,
    col = "black"
  )
  
  segments(
    x0 = x_left, y0 = y_sep,
    x1 = x_right, y1 = y_sep,
    lty = sep_lty, lwd = sep_lwd, col = "black"
  )
  
  text(
    x = x_mid,
    y = y_bottom + bottom_h / 2,
    labels = block2,
    adj = c(0.5, 0.5),
    cex = cex_text,
    family = family,
    col = "black"
  )
  
  if (return_recorded) {
    recordPlot()
  } else {
    invisible(NULL)
  }
}
# Custom function to display one or two exemplar quotes in a boxed panel.
# I wrote this because the original helper plotting functions were unavailable,
# and I needed a way to approximate the visual layout used in the paper.

make_quote_panel <- function(
    quote1,
    quote2 = NULL,
    wrap_width = 35,
    rect_xleft = 0.08,
    rect_ybottom = 0.15,
    rect_xright = 0.92,
    rect_ytop = 0.85,
    padding_x = 0.05,
    padding_y = 0.06,
    sep_gap = 0.05,      # fraction of panel height
    cex_text = 1.1,
    lineheight = 1.2,
    family = "sans",
    border_lwd = 2,
    sep_lwd = 1.5,
    sep_lty = 2,
    add_quotes = FALSE,
    return_recorded = FALSE
) {
  if (length(wrap_width) == 1L) {
    wrap_width <- c(wrap_width, wrap_width)
  }
  if (length(wrap_width) != 2L) {
    stop("wrap_width must have length 1 or 2.")
  }
  if (is.null(quote1) || !nzchar(quote1)) {
    stop("quote1 must be a non-empty string.")
  }
  
  format_quote <- function(x, width) {
    if (is.null(x) || !nzchar(x)) return(NULL)
    x <- if (add_quotes) paste0("\"", x, "\"") else x
    paste(strwrap(x, width = width), collapse = "\n")
  }
  
  q1 <- format_quote(quote1, wrap_width[1])
  q2 <- format_quote(quote2, wrap_width[2])
  
  op <- par(
    mar = c(0, 0, 0, 0),
    xpd = NA,
    family = family,
    lheight = lineheight
  )
  on.exit(par(op), add = TRUE)
  
  plot.new()
  plot.window(xlim = c(0, 1), ylim = c(0, 1), xaxs = "i", yaxs = "i")
  
  # Outer panel
  rect(rect_xleft, rect_ybottom, rect_xright, rect_ytop,
       col = "white", border = "black", lwd = border_lwd)
  
  panel_w <- rect_xright - rect_xleft
  panel_h <- rect_ytop - rect_ybottom
  
  x_left   <- rect_xleft   + padding_x * panel_w
  x_right  <- rect_xright  - padding_x * panel_w
  y_bottom <- rect_ybottom + padding_y * panel_h
  y_top    <- rect_ytop    - padding_y * panel_h
  x_mid    <- (x_left + x_right) / 2
  
  # Single-quote case
  if (is.null(q2)) {
    text(
      x = x_mid,
      y = (y_bottom + y_top) / 2,
      labels = q1,
      adj = c(0.5, 0.5),
      cex = cex_text,
      family = family,
      col = "black"
    )
    
    if (return_recorded) {
      return(recordPlot())
    } else {
      return(invisible(NULL))
    }
  }
  
  # Two-quote case: measure wrapped heights on the active device
  h1 <- strheight(q1, units = "user", cex = cex_text, family = family)
  h2 <- strheight(q2, units = "user", cex = cex_text, family = family)
  
  if (!is.finite(h1) || h1 <= 0) h1 <- 1
  if (!is.finite(h2) || h2 <= 0) h2 <- 1
  
  sep_gap_user <- sep_gap * panel_h
  inner_h <- y_top - y_bottom
  usable_h <- inner_h - sep_gap_user
  
  if (usable_h <= 0) {
    stop("Not enough vertical space after padding and separator gap.")
  }
  
  share1 <- h1 / (h1 + h2)
  share2 <- h2 / (h1 + h2)
  
  top_h <- usable_h * share1
  bottom_h <- usable_h * share2
  
  y_sep <- y_bottom + bottom_h + sep_gap_user / 2
  
  # Top quote
  text(
    x = x_mid,
    y = y_top - top_h / 2,
    labels = q1,
    adj = c(0.5, 0.5),
    cex = cex_text,
    family = family,
    col = "black"
  )
  
  # Dashed separator
  segments(
    x0 = x_left, y0 = y_sep,
    x1 = x_right, y1 = y_sep,
    lty = sep_lty, lwd = sep_lwd, col = "black"
  )
  
  # Bottom quote
  text(
    x = x_mid,
    y = y_bottom + bottom_h / 2,
    labels = q2,
    adj = c(0.5, 0.5),
    cex = cex_text,
    family = family,
    col = "black"
  )
  
  if (return_recorded) {
    recordPlot()
  } else {
    invisible(NULL)
  }
}

#Making Figure 4
jpeg("stm_two_topics_fig4.jpg", width = 8, height = 4, units = "in", res = 500)
make_stm_topic_panel(
  model = mod, # stm object
  topic1 = 2,  # topic number 1
  topic2 = 1, # topic number 2
  n = 8,
  wrap_width = c(70, 70),
  family = "sans"
)
dev.off() #Close the active graphics device so the saved plot is written properly.


```


#Figure 5: Representative Response from Topic
```{r}
#Figure 5 displays exemplar responses from one recovered topic
# to show the kind of text captured by that topic.

#Making Figure 5
quote_t2 <- thoughts$docs[[2]][3]
quote_t2a <- thoughts$docs[[2]][4]
thoughts$docs[[2]]

jpeg("two_quote_panel_top1.jpg", width = 8, height = 4, units = "in", res = 300)
make_quote_panel(
  quote1 = quote_t2,
  quote2 = quote_t2a,
  wrap_width = c(45, 45),
  family = "sans"
)
dev.off() # Close the active graphics device so the saved plot is written properly.
```


#Figure 6: A Representative Response from Topic

```{r}
# Figure 6 displays exemplar responses from the second recovered topic
# so the topic labels can be compared to actual underlying documents.

#Making Figure 6
quote_t1 <- thoughts$docs[[1]][2]
quote_t1a <- thoughts$docs[[1]][5]
thoughts$docs[1]
jpeg("two_quote_panel_top2.jpg", width = 8, height = 4, units = "in", res = 300)
make_quote_panel(
  quote1 = quote_t1,
  quote2 = quote_t1a,
  wrap_width = c(45, 45),
  family = "sans"
)
dev.off() # Close the active graphics device so the saved plot is written properly.
```


#Figure 7: Words and Treatment Effects Associated with Topic 1

```{r}
# Figure 7 estimates post-STM treatment effects on topic prevalence.
# estimateEffect() takes the fitted STM and calculates how the expected proportion
# of each topic changes across the treatment and control conditions.

prep <- estimateEffect(   # Estimate how document-level covariates relate to topic prevalence after fitting the STM.
  c(2, 1) ~ treatment,
  stmobj = mod,
  metadata = meta,
  uncertainty = "Global",
  documents = documents
)

plot( # Plot the estimated difference in topic proportions between treated and control respondents.
  prep,
  covariate = "treatment",
  topics = c(2, 1),
  method = "difference",
  cov.value1 = 1,
  cov.value2 = 0,
  xlab = "Difference in Topic Proportions (Treated-Control)",
  xlim = c(-0.4, 0.3),
  ci.level = 0.90,
  labeltype = "numbers"
)
# Display the top words for the plotted topics so the treatment-effect figure
# can be read together with the substantive content of each topic.
labelTopics(mod, topics = c(2, 1), n = 7)
```
#Figure 8: Party Identification, Treatment, and the Predicted Proportion in Topic
```{r}
# Figure 8 estimates whether the treatment effect on topic prevalence changes across party identification.
# Here, estimateEffect() is a post-estimation step: the STM has already been fit above,
# and this function now estimates the conditional relationship between topic prevalence,
# treatment status, and party ID.

# This chunk is kept together because the second plot is layered on top of the first
# using add = TRUE, so running lines separately can break the final figure.

prep_int <- estimateEffect(
  c(2) ~ treatment * s(pid_rep),
  stmobj = mod,
  metadata = meta,
  uncertainty = "Global",
  documents = documents)

plot(   # Plot the estimated topic prevalence curve for treated respondents across party ID.
  prep_int,
  covariate = "pid_rep",
  topics = 2,
  method = "continuous",
  moderator = "treatment",
  moderator.value = 1,
  ci.level = 0.90,
  xlab = "Party ID",
  ylab = "Mean Topic Proportions",
  xlim = c(0, 1),
  ylim = c(0, 0.7),
  printlegend = FALSE,
  xaxt = "n",
  linecol = "red"
)

plot(
  prep_int,
  covariate = "pid_rep",
  topics = 2,
  method = "continuous",
  moderator = "treatment",
  moderator.value = 0,
  ci.level = 0.90,
  xlab = "Party ID",
  ylab = "Mean Topic Proportions",
  xlim = c(0, 1),
  ylim = c(0, 0.7),
  printlegend = FALSE,
  xaxt = "n",
  linecol = "blue",
  add = TRUE
)

axis(
  1,
  at = c(0, 0.5, 1),
  labels = c("Strong \n Democrat", "Moderate", "Strong \n Republican"),
  padj = 0.5
)

text(0.35, 0.55, "Treated", col = "red")
text(0.75, 0.15, "Control", col = "blue")



# Alternate manual loess version of Figure 8.
# This does not re-estimate the STM. Instead, it smooths the recovered topic proportions
# from the selected STM run in order to mimic the visual shape of the published figure more closely.

# The alternate loess version is included as a visual comparison, not as a separate STM estimate.

topic_k <- 2

treated <- meta$treatment == 1
control <- meta$treatment == 0

y_treat <- mod$theta[treated, topic_k]
x_treat <- meta$pid_rep[treated]

y_ctrl <- mod$theta[control, topic_k]
x_ctrl <- meta$pid_rep[control]

grid <- seq(0, 1, length.out = 200)

fit_treat <- loess(y_treat ~ x_treat, span = 1)
fit_ctrl  <- loess(y_ctrl  ~ x_ctrl,  span = 1)

pred_treat <- predict(fit_treat, newdata = data.frame(x_treat = grid), se = TRUE)
pred_ctrl  <- predict(fit_ctrl,  newdata = data.frame(x_ctrl  = grid), se = TRUE)

plot(grid, pred_treat$fit, type = "l",
     ylim = c(0, 0.7), xlim = c(0, 1),
     xlab = "Party ID", ylab = "Mean Topic Proportions",
     xaxt = "n", col = "red", lwd = 2)

lines(grid, pred_treat$fit + 1.645 * pred_treat$se.fit, col = "red", lty = 2)
lines(grid, pred_treat$fit - 1.645 * pred_treat$se.fit, col = "red", lty = 2)

lines(grid, pred_ctrl$fit, col = "blue", lwd = 2)
lines(grid, pred_ctrl$fit + 1.645 * pred_ctrl$se.fit, col = "blue", lty = 2)
lines(grid, pred_ctrl$fit - 1.645 * pred_ctrl$se.fit, col = "blue", lty = 2)

axis(1, at = c(0, 0.5, 1),
     labels = c("Strong \n Democrat", "Moderate", "Strong \n Republican"))

text(0.35, 0.52, "Treated", col = "red")
text(0.72, 0.16, "Control", col = "blue")

```




#Figures 9 and 10: Fearful Response with High and Low Topic 1
```{r Figures 9 and 10}
# Figures 9 and 10 present representative quotes for high- and low-topic responses.
# I first inspected several candidate quotes from the topic pools. Because the topic numbering
# in my replication did not map cleanly onto the paper, I compared multiple candidates and then
# reordered the search to select quotes that best matched the original figures.


thoughts <- findThoughts(mod, texts=meta$open.ended.response, topics= c(1,2,3), n=10) # Retrieve exemplar documents that load highly on each topic so the replication can show representative quotes.

# Figure 9
# Initial inspection step: check candidate quotes and associated coder labels before selecting the final exemplars.
thoughts$docs[[1]]
meta$fear_ra1[thoughts$index[[1]]]  #In my replication, this recovered topic aligns more closely with paper Topic 2.
meta$fear_ra2[thoughts$index[[1]]]
plotQuote(thoughts$docs[[1]][1], width = 7) # Use stm's quote-plotting helper to display a selected exemplar response.
plotQuote(thoughts$docs[[1]][4], width = 5) # Use stm's quote-plotting helper to display a selected exemplar response.



#Figure10
thoughts$docs[[2]]
meta$fear_ra1[thoughts$index[[2]]] #In my replication, this recovered topic aligns more closely with paper Topic 1.
meta$fear_ra2[thoughts$index[[2]]]
plotQuote(meta$open.ended.response[thoughts$index[[2]][3]], width = 7) # Use stm's quote-plotting helper to display a selected exemplar response.

# Representative quotes in the order that best matches the paper's figures
# thoughts$docs[[2]] = Figure 9 pool (paper Topic 1: jobs/crime/services)
# thoughts$docs[[1]] = Figure 10 pool (border/legality/amnesty contrast)

thoughts <- findThoughts(  # Retrieve exemplar documents that load highly on each topic so the replication can show representative quotes.
  mod,
  texts = meta$open.ended.response,
  topics = c(2, 1, 3),
  n = 10
)

# Inspect coder labels and candidate quotes
meta$fear_ra1[thoughts$index[[1]]]
meta$fear_ra2[thoughts$index[[1]]]
thoughts$docs[[1]]

meta$fear_ra1[thoughts$index[[2]]]
meta$fear_ra2[thoughts$index[[2]]]
thoughts$docs[[2]]

# Final selections
# Figure 9 analogue: selected for substantive similarity to the paper's low-topic example,
# even though the replicated topic mapping is not identical.
plotQuote(thoughts$docs[[1]][3], width = 7) # Use stm's quote-plotting helper to display a selected exemplar response.
# Figure 10 analogue: high predicted proportion of paper Topic 1
plotQuote(thoughts$docs[[2]][6], width = 35) # Use stm's quote-plotting helper to display a selected exemplar response.

```



#Cleaning for Rand Analysis: Public Goods Games Experiment
```{r}
# This chunk attempts to import and repair the RAND replication files, including vocabulary alignment issues. 
dataldac <- readLdac("final_tdm.csv") # Read the RAND public goods game document-term matrix.
#Testing the File Pairing
term_ids <- unlist(lapply(dataldac, function(x) x[1, ])) # Check whether the term IDs in the document-term matrix line up with the supplied vocabulary.
max(term_ids)
length(vocabcoarse)
range(term_ids)

#Inspecting the vocab files
readLines("tdm89_vocab.csv", n = 10) # Inspect the vocabulary file directly to diagnose the mismatch.
tail(readLines("tdm89_vocab.csv"), 10)


# original vocab 
vocabcoarse <- read.csv("tdm89_vocab.csv")$term #Load the RAND vocabulary file.

# Add placeholder terms for the missing IDs so prepDocuments() can at least attempt
# to align the document-term matrix and the vocabulary to the same length.
missing_ids <- 474:493 
vocab_padded <- c(vocabcoarse, paste0("MISSING_TERM_", missing_ids))

length(vocab_padded)   # should now be 494


# Shift document term IDs to 1-indexing for stm/prepDocuments compatibility.
dataldac_fixed <- lapply(dataldac, function(doc) {
  rbind(doc[1, ] + 1L, doc[2, ])
})

# Attempt to prepare the RAND documents after padding the vocabulary and shifting IDs.
meta <- read.csv("tdm89_metadata.csv")
meta$treatment <- as.numeric(meta$condition_number==9)
data <- prepDocuments(dataldac_fixed, vocab_padded, meta = meta)
documents <- data$documents
vocab <- data$vocab
meta <- data$meta

# Check whether the treatment variable remains usable after document preparation.
sum(is.na(meta$treatment))
sum(!is.na(meta$treatment))
table(meta$treatment, useNA = "ifany")
```
In the end, this workaround was not successful. 
The metadata, vocabulary, and treatment files in the replication package did 
not align cleanly enough to construct a valid STM input. Because of those 
mismatches, I was unable to prepare the data in a way that allowed the model to be estimated.






```{r, include = FALSE, SettingUptheModel}
# Hidden troubleshooting chunk: attempted RAND model recovery after the file repairs above.
# I kept this chunk to document the failed modeling attempt, even though it was not retained
# as part of the final replication because the input files remained inconsistent.

#set.seed(54321) #selecting the seed
#
#models <- selectModel(documents, vocab, 5, prevalence= ~meta$treatment, runs=50)
#plotModels(models) # Plot diagnostic information for the candidate models to help choose the best-fitting/interpretable run.
#length(documents)
#nrow(meta)
#table(meta$treatment, useNA = "ifany")
#
#mod <- models$runout[[3]]
##making the model
#mod <- stm(documents, vocab, 5, prevalence=~meta$treatment, seed=1601900)
#labeltopics(mod, n=24) #pulling the topic labels

```
#Set Up Attempt for the Rand Experiment
```{r, include = FALSE}
# This section attempts to prepare the alternate RAND metadata/vocabulary pairing used for later figures.
#dataldac <- readLdac("final_tdm.csv")
#vocabcoarse <- read.csv("tdm1011_vocab.csv")$term
#meta3 <- read.csv("tdm1011_metadata.csv")
#meta3$treatment <- as.numeric(meta3$condition_number==10)
#
#data <- prepDocuments(dataldac, vocabcoarse, meta=meta3)
#documents <- data$documents
#vocab <- data$vocab
#meta3 <- data$meta
#
## Select a candidate STM for the second RAND experiment.
#set.seed(01238)
#models <- selectModel(documents, vocab, 5, prevalence=~meta3$treatment, runs=50)
#plotModels(models)
#mod <- models$runout[[4]]
#
## Recover the selected model directly using the saved seed from the chosen run.
##seed: 4671639
#mod <- stm(documents, vocab, 5, prevalence=~meta$treatment, seed=4671649)
#labeltopics(mod)
```






#Figure 11
```{r}
# This figure was intended to show the RAND STM treatment effect,
# but it could not be reproduced because the RAND model could not be validly estimated.
#Treatment effect plot
#prep <- prep.plot(mod, meta$treatment, theta.uncertainty="Global", documents=documents, ci.level=.9)
#ploteffect(prep,choosetopics=c(1,4),model=mod, xlab="Difference in Topic Proportions (Treated-Control)", labeltype="numbers", xlim=c(-.2, .15))
#plottopics(mod, topics=c(1,4),width=50)
```


#Figure 12
```{r}
# This chunk plots the corresponding treatment effect from the LDA comparison model.
#Treatment effect plot
#prep <- prep.plot(mod, meta$treatment, theta.uncertainty="Global", documents=documents, ci.level=.9)
#ploteffect(prep, choosetopics=c(1,5), model=mod, xlab="Difference in Topic Proportions (Treated-Control)", labeltype="numbers", xlim=c(-.18,.15))
#plottopics(mod, topics=c(1,5), width=50)
```



#Figure 13
```{r}
# This chunk produces the normalized contribution plot used to compare topic content and treatment effects.
#Normalized Contributon Plot
#brandonexample <- prep.plot(mod, meta$normalized.contribution, topics=c(1,4),theta.uncertainty="Local", span=2, ci.level=.9, documents=documents, dv="covariate")
#prep2 <- prep.plot(mod, meta$normalized.contribution, topics=c(1,4),theta.uncertainty="Global", span=2, ci.level=.9, documents=documents, dv="covariate")
#ploteffect(prep2, choosetopics=c(1,2),model=mod, xlab="Mean Topic Proportions", ylab="Mean Normalized Contributions", labeltype="numbers", xlim=c(.05,.6), ylim=c(.15,1), #linecol=c("blue", "red"))


```
#Figures 15 and 16
```{r}

# This chunk fits a content-covariate STM and generates the gender-comparison figures from the RAND analysis.
#Select model
#set.seed(02138)
#models <- selectModel(documents, vocab, 5, prevalence=~meta$treatment, content=~meta$female, runs=50)
#plotModels(models)
#mod2 <- models$runout[[9]]
#mod$settings$seed

#Recover selected model
#Seed 4005847
#meta$gender <- ifelse(meta$female==1, "female", "male")
#mod2 <- stm(documents, vocab, 5, prevalence=~meta$treatment, content=~meta$gender, seed=4005847)
#labeltopics(mod2)
#
#prep <- prep.plot(mod2, meta$treatment, theta.uncertainty="None", documents=documents, ci.level=.95)
#ploteffect(prep,model=mod2, xlab="Treated-Control", labeltype="numbers")
#
#plotPerspectives(mod2, 3, M=50, labels=c("Female", "Male"))
#
#plottopics(mod2, 3, width=20, covariatelabs=c("Female", "Male"), n=40)

```

##ANES Analysis

#ANES Set Up
```{r, include = FALSE}
# This section replicates the ANES-based portion of the paper by comparing
# hand-coded issue categories to STM-estimated topic prevalence.

# Keep only respondents with usable party ID, education, and age values,
# matching the variables needed for the prevalence specification below.
dataldac <- readLdac("final_anes.csv")
vocabcoarse <- read.csv("final_anes_vocab.csv")$term
meta4 <- read.csv("final_anes_metadata.csv")

dataldac <- dataldac[meta4$pid_summary>0 & meta4$highest.grade.completed>0 & meta4$age>0]
meta4 <- meta4[meta4$pid_summary>0 & meta4$highest.grade.completed>0 & meta4$age>0,]

data <- prepDocuments(dataldac, vocabcoarse, meta=meta4) # Clean and align the ANES documents, vocabulary, and metadata for STM.
documents <- data$documents # Extract the processed document list used by stm.
vocab <- data$vocab # Extract the cleaned vocabulary matched to the processed documents.
meta4 <- data$meta # Keep the metadata rows that remain after document preparation.

#Select and Configure Model
set.seed(743)

models <- selectModel( documents,vocab, 60, prevalence = ~ s(pid_summary) + s(age) + s(highest.grade.completed),
  data = meta4,runs = 50)
plotModels(models) # Compare candidate ANES models across STM diagnostics before selecting one for replication.
mod5 <- models$runout[[5]] # Select the candidate model used for the later ANES figures.
labelTopics(mod5, n = 7) # Inspect top words to identify the substantive content of recovered topics.
```

#Figure 17
```{r, fig.height=15, fig.width=10}
## Figure 17 displays the top ANES topics using STM's built-in topic plot,
# providing an overview of the recovered topic structure.

mod <- models$runout[[5]]
jpeg("ANES_topics.jpg", width = 10, height = 15, units = "in", res = 300)
plot.STM(mod, n=3, threshold=3, numtopics=50, labeltype="frex")
dev.off()
```


#Figure 18 
```{r}
# Figure 18 compares two approaches within the same substantive topic.
# The left panel recreates the hand-coded war-topic relationship from the paper,
# and the right panel compares it to the analogous STM-based interaction estimate.
# This comparison shows whether the STM-based topic behaves similarly
# to the hand-coded war category used in the original analysis.
par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
# Original code uses ANES category 3 for the plotted "war/Iraq War" indicator

#ANES Hand-coded Data

anes <- 3

meta4$warall <- ifelse(
  meta4$mippol1_code1 == anes |
    meta4$mippol1_code2 == anes |
    meta4$mippol1_code3 == anes |
    meta4$mippol1_code4 == anes |
    meta4$mippol1_code5 == anes |
    meta4$mippol1_code6 == anes |
    meta4$mippol1_code7 == anes |
    meta4$mippol1_code8 == anes,
  1, NA)
meta4$warall[is.na(meta4$warall)] <- 0

main <- "ANES War Topic and Education"

# Democrats: pid_summary < 3
sub_dem <- subset(
  meta4,
  pid_summary < 3 & !is.na(highest.grade.completed) & !is.na(warall)
)

fit_dem <- loess(warall ~ highest.grade.completed, data = sub_dem, span = 2)

newdat_dem <- data.frame(
  highest.grade.completed = seq(13, 17, length.out = 100)
)

pred_dem <- predict(fit_dem, newdata = newdat_dem, se = TRUE)

plot(
  pred_dem$fit ~ newdat_dem$highest.grade.completed,
  type = "l",
  ylim = c(0, 0.1),
  xlim = c(13, 17),
  col = "blue",
  xlab = "Years of Education",
  ylab = "Probability of Topic",
  main = main
)

lines(
  (pred_dem$fit + 1.64 * pred_dem$se) ~ newdat_dem$highest.grade.completed,
  col = "blue", lty = 2
)
lines(
  (pred_dem$fit - 1.64 * pred_dem$se) ~ newdat_dem$highest.grade.completed,
  col = "blue", lty = 2
)

# Republicans: pid_summary > 3
sub_rep <- subset(
  meta4,
  pid_summary > 3 & !is.na(highest.grade.completed) & !is.na(warall)
)

fit_rep <- loess(warall ~ highest.grade.completed, data = sub_rep, span = 2)

newdat_rep <- data.frame(
  highest.grade.completed = seq(13, 17, length.out = 100)
)

pred_rep <- predict(fit_rep, newdata = newdat_rep, se = TRUE)

lines(
  pred_rep$fit ~ newdat_rep$highest.grade.completed,
  type = "l",
  col = "red"
)

lines(
  (pred_rep$fit + 1.64 * pred_rep$se) ~ newdat_rep$highest.grade.completed,
  col = "red", lty = 2
)
lines(
  (pred_rep$fit - 1.64 * pred_rep$se) ~ newdat_rep$highest.grade.completed,
  col = "red", lty = 2
)

text(14, 0.02, "Republican", col = "red", cex = 1.2)
text(14.5, 0.08, "Democrat", col = "blue", cex = 1.2)

# STM-based interaction plot for education and the recovered war topic.
# This estimateEffect() model asks whether the relationship between education and topic prevalence
# varies across partisan groups in the STM output.

meta4$pid_group <- ifelse(
  meta4$pid_summary < 3, "Democrat",
  ifelse(meta4$pid_summary > 3, "Republican", "Middle")
)

meta4$pid_group <- factor(
  meta4$pid_group,
  levels = c("Republican", "Middle", "Democrat")
)

table(meta4$pid_group, useNA = "ifany")

prep_ed <- estimateEffect(  # Estimate how document-level covariates relate to topic prevalence after fitting the STM.
  c(23) ~ highest.grade.completed * pid_group,
  stmobj = mod5,
  metadata = meta4,
  uncertainty = "Global",
  documents = documents
)

# Democrats
plot(
  prep_ed,
  covariate = "highest.grade.completed",
  topics = 23,
  method = "continuous",
  moderator = "pid_group",
  moderator.value = "Democrat",
  ci.level = 0.90,
  ylab = "Probability of Topic",
  xlab = "Years of Education",
  ylim = c(0, 0.1),
  xlim = c(13, 17),
  printlegend = FALSE,
  linecol = "blue",
  main = "STM War Topic and Education"
)

# Republicans
plot(
  prep_ed,
  covariate = "highest.grade.completed",
  topics = 23,
  method = "continuous",
  moderator = "pid_group",
  moderator.value = "Republican",
  ci.level = 0.90,
  ylab = "Probability of Topic",
  xlab = "Years of Education",
  ylim = c(0, 0.1),
  xlim = c(13, 17),
  printlegend = FALSE,
  linecol = "red",
  add = TRUE
)

text(16, 0.09, "Democrat", col = "blue", cex = 1.2)
text(14.5, 0.03, "Republican", col = "red", cex = 1.2)

```


